A Simple, Useful, and Only Slightly Wrong Model of Virtual Memory

### 2010-10-06 article.tex
11 pt font.  Looks uglier, conforms to guidelines.

\documentclass{acmtrans2m}

\def\endbottomstuff{\end{figure}\gdef\permission{}}

\def\runningfoot{\def\@runningfoot{}}
\def\firstfoot{\def\@firstfoot{}}

\usepackage{url}
\usepackage{salgorithm}
\usepackage{graphicx} \graphicspath{{sorts/}}
\usepackage{fancyvrb}

%\acmVolume{V} 
%\acmNumber{N} 
%\acmYear{YY} 
%\acmMonth{M} 
\markboth{Peter Boothe}{A Simple, Useful Model of Virtual Memory} 
\title{A Simple, Useful, and Only Slightly Wrong Model of Virtual Memory}
\author{PETER BOOTHE\\
Manhattan College
}
\begin{abstract} 
We define a simple model of virtual memory in an effort to predict the number
of page faults that an algorithm will incur over the course of its lifetime.
We then use that model to analyze how many page faults we would expect to see
with two of the best-known and most well-studied sorting algorithms.  Then, we
run multiple experiments in an effort to show that the predictions of our model
match up with reality.
\end{abstract}

\category{F.2.0}{Analysis of Algorithms and Problem Complexity}{General} 
\category{D.4.2}{Operating Systems}{Storage Management} [Virtual Memory]
\category{D.4.8}{Operating Systems}{Performance} [Modeling and Prediction]
\terms{Algorithms, Measurement, Experimentation} 
\keywords{Virtual Memory, Algorithm Analysis}

@@

\begin{document} 
\begin{bottomstuff} 
This research was supported by a summer grant from Manhattan College.\\
Author's Address: \url{peter.boothe@manhattan.edu}, or P. Boothe, Manhattan College Math and Computer Science Dept. RLC 201, 3840 Corlear Ave, Bronx, NY 10463
\end{bottomstuff} 
\maketitle

@@

Our analysis of \heapsort\ is the easiest.  The \heapsort\ algorithm, which we
take from \citeN{clrs}, is defined, with all of its supporting functions, in
Figure~\ref{fig:hsort}.  The implementation we tested is from the BSD standard
library, which is available as {\tt libbsd} on Linux.

\begin{narrowfig}{1.5in}
\begin{algorithm}
\Algorithm{HeapSort}$(array, size)$\+\\
    Build-Max-Heap$(array, size)$\\
    \For $i \gets size$ \DownTo\ $2$\+\\
        exchange $array[1] \leftrightarrow array[i]$\\
        $size \gets size - 1$\\
        Max-Heapify$(array, 1, size)$\-\\
    \Return $array$\-\\
\\
\Algorithm{Build-Max-Heap}$(array, size)$\+\\
    \For $i \gets \lfloor \frac{size}{2} \rfloor$ \DownTo\ $1$\+\\
        Max-Heapify$(array, i, size)$\-\-\\
\\
\Algorithm{Max-Heapify}$(array, i, size)$\+\\
    $l \gets 2*i$\\
    $r \gets 2*i + 1$\\
    $largest \gets i$\\
    \If $l < size$ and $array[l] > array[largest]$\+\\
            $largest \gets l$\-\\
    \If $r < size$ and $array[r] > array[largest]$\+\\
            $largest \gets r$\-\\
    \If $largest \neq i$\+\\
        exchange $array[i] \leftrightarrow array[largest]$\\
        Max-Heapify$(array, largest, size)$
\end{algorithm}

\caption{\heapsort\ pseudo code.}
\label{fig:hsort}
\end{narrowfig}

@@

The \mergesort\ algorithm we adapt from \citeN{clrs} with the memory-efficiency
improvement that, in the merge step, we only copy out the first half of the
array, and merge the second half in-place.  Our implementation may be found in
Figure~\ref{fig:msort}.  In our analysis of page faults, the only consideration
is the number and location of memory accesses, and memory accesses only occur
in the ``merge'' portion of the algorithm.

@@

\begin{narrowfig}{3in}
\begin{Verbatim}[numbers=left,numbersep=3pt,xleftmargin=20pt]
#include <stdio.h>
#include <stdlib.h>

@@

        return 0;
}
\end{Verbatim}
\caption{Our test harness for both algorithms}
\label{fig:harness}
\end{narrowfig}
\subsection{Experimental Setup}

@@

\bibliographystyle{acmtrans}
\bibliography{bibliography}

@@

\begin{received} 
Submission to ALENEX 2011
\end{received}
\end{document}

>>>>

\documentclass[11pt]{article}

\usepackage{url}
\usepackage{salgorithm}
\usepackage{graphicx} \graphicspath{{sorts/}}
\usepackage{fancyvrb}
\usepackage[letterpaper]{geometry}

%\acmVolume{V} 
%\acmNumber{N} 
%\acmYear{YY} 
%\acmMonth{M} 
%
%\category{F.2.0}{Analysis of Algorithms and Problem Complexity}{General} 
%\category{D.4.2}{Operating Systems}{Storage Management} [Virtual Memory]
%\category{D.4.8}{Operating Systems}{Performance} [Modeling and Prediction]
%\terms{Algorithms, Measurement, Experimentation} 
%\keywords{Virtual Memory, Algorithm Analysis} 
\markboth{Peter Boothe}{A Simple, Useful Model of Virtual Memory} 
\title{A Simple, Useful, and Only Slightly Wrong Model of Virtual Memory}
\author{Peter Boothe\thanks{This research was supported by a summer grant from Manhattan College.
Author's Address: \texttt{peter.boothe@manhattan.edu}, or P. Boothe, Manhattan College Math and Computer Science Dept. RLC 201, 3840 Corlear Ave, Bronx, NY 10463
}\\
Manhattan College
}

@@

\begin{document}

\maketitle

\begin{abstract} 
We define a simple model of virtual memory in an effort to predict the number
of page faults that an algorithm will incur over the course of its lifetime.
We then use that model to analyze how many page faults we would expect to see
with two of the best-known and most well-studied sorting algorithms.  Then, we
run multiple experiments in an effort to show that the predictions of our model
match up with reality.
\end{abstract}

@@

Our analysis of \heapsort\ is the easiest.  The \heapsort\ algorithm, which we
take from \cite{clrs}, is defined, with all of its supporting functions, in
Figure~\ref{fig:hsort}.  The implementation we tested is from the BSD standard
library, which is available as {\tt libbsd} on Linux.

\begin{figure}
\begin{algorithm}
\Algorithm{HeapSort}$(array, size)$\+\\
    Build-Max-Heap$(array, size)$\\
    \For $i \gets size$ \DownTo\ $2$\+\\
        exchange $array[1] \leftrightarrow array[i]$\\
        $size \gets size - 1$\\
        Max-Heapify$(array, 1, size)$\-\\
    \Return $array$\-\\
\\
\Algorithm{Build-Max-Heap}$(array, size)$\+\\
    \For $i \gets \lfloor \frac{size}{2} \rfloor$ \DownTo\ $1$\+\\
        Max-Heapify$(array, i, size)$\-\-\\
\\
\Algorithm{Max-Heapify}$(array, i, size)$\+\\
    $l \gets 2*i$\\
    $r \gets 2*i + 1$\\
    $largest \gets i$\\
    \If $l < size$ and $array[l] > array[largest]$\+\\
            $largest \gets l$\-\\
    \If $r < size$ and $array[r] > array[largest]$\+\\
            $largest \gets r$\-\\
    \If $largest \neq i$\+\\
        exchange $array[i] \leftrightarrow array[largest]$\\
        Max-Heapify$(array, largest, size)$
\end{algorithm}

\caption{\heapsort\ pseudo code.}
\label{fig:hsort}
\end{figure}

@@

The \mergesort\ algorithm we adapt from \cite{clrs} with the memory-efficiency
improvement that, in the merge step, we only copy out the first half of the
array, and merge the second half in-place.  Our implementation may be found in
Figure~\ref{fig:msort}.  In our analysis of page faults, the only consideration
is the number and location of memory accesses, and memory accesses only occur
in the ``merge'' portion of the algorithm.

@@

\begin{figure}
\begin{Verbatim}[numbers=left,numbersep=3pt,xleftmargin=20pt]
#include <stdio.h>
#include <stdlib.h>

@@

        return 0;
}
\end{Verbatim}
\caption{Our test harness for both algorithms}
\label{fig:harness}
\end{figure}
\subsection{Experimental Setup}

@@

\bibliographystyle{abbrv}
\bibliography{bibliography}

@@

%\begin{received} 
Submission to ALENEX 2011
%\end{received}
\end{document}
